---
title: 'A3: Incarceration'
output:
  html_document: default
  pdf_document: default
---

```{r include=F}
source("summary.R")
source("chart1_example.R")
source("chart2_example.R")
source("map_example.R")
```

### Introduction

I investigated the research question concerning the overall Prison Population Rate in Washington (WA) over time, the comparison of Prison Rates between males and females, and the Prison Population across various states in the United States. My selection of these three questions was driven by the desire to understand the larger context, focusing on ethnic disparities in localized areas and overarching trends nationwide. On a more localized scale, I examined the prison population rate among different ethnic groups within 39 counties in Washington state. Conversely, I assessed the total prison population in individual states on a broader scale, evaluating regional variations on a U.S. map.

### Summary Information

My revised and reduced dataset comprises a total of `r summary_info$num_observations` individual observations and `r summary_info$num_column` characteristics or elements under examination. This data encompasses the prison population across `r summary_info$num_states` states within the U.S. The year marked by the peak in prison population numbers is `r summary_info$high_p_year`, and the state registering the most significant number of prison inhabitants is `r summary_info$high_p_state`.

### The Dataset

Who collected the data?  
How was the data collected or generated?  
Why was the data collected?  
How many observations (rows) are in your data?  
How many features (columns) are in the data?  
What, if any, ethical questions or questions of power do you need to consider when working with this data?  
What are possible limitations or problems with this data?   (at least 200 words)


### Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - This may mean showing the same measure for different locations or different racial groups. Think carefully about a meaningful comparison of locations (e.g., the top 10 counties in a state, top 10 states, etc.)
- You must have clear x and y axis labels
- The chart needs a clear title
- You need a legend for your different line colors and a clear legend title

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("chart1_example.R"), warning=F, message=F}
chart1
```

### Variable Comparison Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show how two different (continuous) variables are related to one another. Again, think carefully about what such a comparison means and what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title
- If you choose to add a color encoding (not required), you need a legend for your different color and a clear legend title

``````{r, echo = FALSE, code = readLines("chart2_example.R"), warning=F, message=F}
chart2
```
### Map

This map of the United States illustrates the prison population in each state, with states depicted in darker shades of red having more substantial numbers of inmates. The map provides an overall perspective of the distribution of the prison population across the U.S. From the visual representation, it is clear that Texas leads the nation with the most significant number of incarcerated individuals, bearing the heaviest prison burden, followed by Georgia. Other states display less variation amongst themselves, with the central region tending to have a higher prison population.

```{r, echo = FALSE, warning=F, message=F}
us_map_pop_prison
```