---
title: 'A3: Incarceration'
output:
  html_document: default
  pdf_document: default
---

```{r include=F}
source("summary.R")
source("chart1_example.R")
source("chart2_example.R")
source("map_example.R")
```

### Introduction

I investigated the research question concerning the overall Prison Population Rate in Washington (WA) over time, the comparison of Prison Rates between males and females, and the Prison Population across various states in the United States. My selection of these three questions was driven by the desire to understand the larger context, focusing on ethnic disparities in localized areas and overarching trends nationwide. On a more localized scale, I examined the prison population rate among different ethnic groups within 39 counties in Washington state. Conversely, I assessed the total prison population in individual states on a broader scale, evaluating regional variations on a U.S. map.

### Summary Information

My revised and reduced dataset comprises a total of `r summary_info$num_observations` individual observations and `r summary_info$num_column` characteristics or elements under examination. This data encompasses the prison population across `r summary_info$num_states` states within the U.S. The year marked by the peak in prison population numbers is `r summary_info$high_p_year`, and the state registering the most significant number of prison inhabitants is `r summary_info$high_p_state`.

### The Dataset

The Vera Institute of Justice assembled the incarceration trends dataset, incorporating county-level jail data from 1970 to 2018 and county-level prison data from 1983 to 2016. This data compilation was conducted through two Bureau of Justice Statistics data collections and blends sources from the National Correction Reporting Program, Annual Survey of Jails, Census of Jails, as well as population and admissions variables from the Death in Custody Reporting Program. The information was gathered with the intent of visualization using various data tools. In dataset about prison/jail population per 100,000 people (rate) per county/state, there are 54793 observations and 23 variables. In dataset about priso prison/jail population per 100,000 people (rate) per county/state starting at 1990 in Washington, there are 1131 observations and 6 variables. In dataset about prison population per county/state, there are 153811 observations and 36 variables. There is an acknowledgment that human factors may have led to some unavoidable minor inaccuracies in calculation. However, these errors are deemed negligible for a comprehensive trend analysis. The research and study on prisoners should be approached with caution, honoring the principles of justice and respect for individuals. Moreover, as a continuation of the justice concept, the idea of collaborative responsibility should be woven into a revised ethical framework to foster ethical growth. By actively engaging with all relevant stakeholders and maintaining open communication across concerned areas, researchers can ensure they create ethical environments that promote respect and minimize exploitation in every research context.

### Trends Over Time Chart

This line graph illustrates the prison or jail population rate per 100,000 individuals in Washington, segmented by race, including AAPI, Black, Latinx, Native, and White, starting from 1990. It reveals that the prison population rate for the Black community has been on a downward trend from around 2000 to the present, while the Native population experienced an upward trend with a recent sharp decline. In comparison, the Latinx, AAPI, and White groups have relatively lower prison population rates. Currently, the Native race has the highest prison population rate.

``````{r, echo = FALSE, code = readLines("chart1_example.R"), warning=F, message=F}
chart1
```

### Variable Comparison Chart

The chart analyzes the mean prison rate for both males and females, plotting them on a scatter plot complemented by a linear regression line. This visualization emphasizes a pronounced positive linear relationship between female and male prison rates, demonstrating that an increase in the average prison rate for females coincides with a similar rise in the average prison rate for males.

``````{r, echo = FALSE, code = readLines("chart2_example.R"), warning=F, message=F}
chart2
```

### Map

This map of the United States illustrates the prison population in each state, with states depicted in darker shades of red having more substantial numbers of inmates. The map provides an overall perspective of the distribution of the prison population across the U.S. From the visual representation, it is clear that Texas leads the nation with the most significant number of incarcerated individuals, bearing the heaviest prison burden, followed by Georgia. Other states display less variation amongst themselves, with the central region tending to have a higher prison population.

```{r, echo = FALSE, warning=F, message=F}
us_map_pop_prison
```